\section{Termination}
\subsection{Introduction}
\defe{Kolawski slogan}{The \sub{} states the following:
\begin{quote}
An \inre{Algorithm} is \inre{Logic} and \inre{Control}
\end{quote} Where \inre{Logic} is the declarative semantics and \inre{Control} a aspect yielding efficient proofs where \inre{Termination} is a minimum. \inre{Termination} in general however is undecidable. Basically there are two research lines regarding this problem:
\begin{enumerate}
 \item Necessary and sufficient conditions: manual methods to prove termination of a specific program.
 \item Decidable sufficient conditions: automated systems that can conservatively decide wether a program will terminate.
\end{enumerate} Termination also depends on the query: for some queries a program will terminate while for other it will get stuck in a loop. Therefore there are several forms of termination: \inre{Universal termination}, \inre{Existential termination} and \inre{Strong termination}.}
\lema{Incompletenss results}{Assume there exists a function $\mbox{terminates}:P\rightarrow\BB$, which returns true if the program terminates and false otherwise. Then we can construct a program which calls that function and goes into a loop when the function returns true and stops otherwise. Since the program terminates when the program will not terminate and will not terminate when it would terminate, we have a contradiction.}
\defe{Universal termination}{Universal termination states that the program can find all solutions for a query and terminate.}
\defe{Existential termination}{This form of termination states that the program can at least will finitely fail or find a solution. After the first solution, the program may loop. Universal termination implies existential termination. Proving existential termination in the absence of universal termination however is much harder because of the order of branches in the \inre{SLDNF tree}s on must take into account.}
\defe{Strong termination}{\sub{} states that a program terminates independent of the computational rule. This is a quite rare condition.}
\defe{Termination proof}{A termination proof proves one of the termination conditions for a specific program and a specific query. All proofs are based in \inre{Well-founded ordering}: by defining a partial order on the \inre{Well-founded set}. Sequences with a decreasing set are always finite and thus will terminate. Therefore by taking the states of computation as elements and defining an order on the we can decide about termination in some situations.}
\defe{Well-founded order}{For every two states we define the \sub{} such that: if $s_2$ is the successor of $s_1$, $\fun{f}{s_1}<\fun{f}{s_2}$. If this rule is satisfied then the set of states of the computation is a \inre{Well-founded set} and the computation will terminate.}
\lema{Decreasing sequence}{Each finite computation can be mapped on a decreasing sequence in a \inre{Well-founded ordering}.
\begin{proof}
$s_0,s_1,\ldots,s_n$ is a finite sequence. By defining $\fun{f}{s_i}=n-i$ we have $\fun{f}{s_i}=n-i>n-i-1=\fun{f}{s_{i+1}}$ and $\NN$ is a \inre{Well-founded set}.
\end{proof}}
\lema{Termination of programs without variables}{For programs without variables, one can define the order by de size of the ground atom in $\NN$. If for all clauses \clausea{A}{B}{n}, $\forall i:\funm{size}{A}>\funm{size}{B_i}$, then this is a \inre{Well-founded order} over the nodes of any \inre{SLD-tree}. Note that this rule is only applicable for programs without variables.
\begin{proof}
Define $\fun{f}{\clausez{A}{n}}=\accl{\funm{size}{A_1},\ldots,\funm{size}{A_n}}$\footnote{This is a multiset: each item can occur more than once.} Then we define the order relation as: $\accl{a}\cup S>\accl{b_1,\ldots,b_n}\cup S$ if $\forall i: a > b_i$. Since the resolvent is of $\leftarrow A_1,\ldots,A_i,\ldots,A_n$ is $\leftarrow A_1,\ldots,A_{i-1},B_1,\ldots,B_m,A_{i+1},\ldots,A_n$, we have $\fun{f}{\leftarrow A_1,\ldots,A_i,\ldots,A_n}>\fun{f}{\leftarrow A_1,\ldots,A_{i-1},B_1,\ldots,B_m,A_{i+1},\ldots,A_n}$ thus the program is finite.
\end{proof}}
\note{Problems with the termination of problems with variables}{
One could try to set the size of the predicate by the number of functors and for all ground substitutions ensure that the size of the number of functions goes down. But one can also apply non-terminating queries using variables. For instance $\leftarrow\funm{append}{L_1,[],L_3}$. One resolution step can yield the goal, but also instantiates the next goal. The \inre{Backpropagation effect} of unification therefore can lead to the unbounded increase in size of the original goal. This can be countered by reducing to programs where unification only happens in one way: instantiating variables in the head but not in the original goal. In that case the ``size'' looks only at parts of terms of calls which are not further initiated. Another difficulty are the variables local to the body. When one calls the head with the first argument ground, the size cannot consider the second argument due to bacpropagation. Therefore there is no size possible such that the size for the head can be an upperbound for every grounded substitution in the body. Therefore one will need to find a relationship between the size of the variables in the head and the size of the variables in the body.
}
\subsection{Termination for definite programs}
\subsubsection{Strong termination}
\defe{Level mapping}{A \sub{} maps all the elements of the \inre{Herbrand universe} to the \inre{Natural numbers} thus $f:B_P\rightarrow\NN$.}
\defe{Recurrent program}{A program is a \sub{} if and only if there exist a \inre{Level mapping} $f$ such that for every $A\leftarrow B_1,\ldots,B_n\in\funm{ground}{P}$, $\forall i:\fun{f}{A}>\fun{f}{B_i}$.}
\lema{Recurrent programs terminate}{$P$ is \inre{Recurrent} if and only if $P$ is terminating for all ground queries $\leftarrow A$. Therefore it is independent of the computational rule and thus \inre{Strong termination}.
\begin{proof}
\begin{description}
 \item [$\Rightarrow$] Assume $P$ terminates for all ground queries $\leftarrow A$, let $\fun{f}{A}$ be the length of the longest derivation in the SLD-tree of $\leftarrow A$. For each ground rule $A\leftarrow B_1,\ldots,B_n$, the longest derivation of $\leftarrow A$ is greater than the longest derivation of $\leftarrow B_i$ and therefore $P_i$ is recurrent.
 \item [$\Leftarrow$] Assume $P$ is recurrent. Assume there is an infinite derivation $\leftarrow A,\ \leftarrow G_1,\ \leftarrow G_2,\ \ldots$. Let $\leftarrow G_i\theta=\leftarrow A_1\theta,\ldots,A_i\theta,\ldots,A_m\theta$ and let $\leftarrow G_{i+1}\theta=\leftarrow A_1\theta,\ldots,A_{i-1}\theta,B_1\theta,\ldots,B_k\theta,A_{i+1}\theta,\ldots,A_m\theta$. Further we define the size of $\leftarrow G_1\theta$ as $\accl{\fun{f}{A_1\theta},\ldots,\fun{f}{A_i\theta},\ldots,\fun{f}{A_m\theta}}$. Then $\funm{size}{\leftarrow G_i\theta}>\funm{size}{\leftarrow G_{i+1}\theta}$ according to order relations over the multisets, thus the successive states in the derivation are a \inre{Well-founded set} and the derivation is finite. This lead to a contradiction of the infinite derivation.
\end{description}
\end{proof}}
\defe[Boundedness]{Bounded atom}{Let $f$ be a \inre{Level mapping}. An atom $A$ is bounded with respect to $f$ if $\accl{\fun{f}{A\theta}|A\theta\mbox{ is ground}}$ is bounded in $\NN$. Therefore the predicate has an upper bound and back propagation cannot infinitely increase the size of the original query.}
\lema{Termination of bounded queries}{If $P$ is \inre{Recurrent} with respect to a level mapping $f$, then $P$ is strongly terminating for all atomic queries which are bounded with respect to $f$.
\begin{proof}
Assume a bounded query for which an infinite derivation exists. By applying a ground substitution we get a bounded size which contradicts the lemma an recurrency of $P$.
\end{proof}}
\lema{Termination of atomic queries}{If atomic queries bounded with respect to $f$ always strongly terminate, then $P$ is recurrent with respect to some \inre{Level mapping} $g$.
\begin{proof}
All ground queries are bounded with respect to $f$ hence strongly terminate. Hence $P$ is recurrent with some level mapping (according to the previous lemma).
\end{proof}}
\subsubsection{LD-termination}
\defe{Left termination}{\sub{} is a form of termination where the resolution phase always selects the \inre{Leftmost atom}. All ``good'' programs are left terminating.}
\defe{Acceptable program}{A program $P$ is acceptable if there exists a level mapping $f$ and a model $I$ such that for each clause $A\leftarrow B_1,\ldots,B_k\in\funm{ground}{P}$, if $I\vDash B_1,\ldots, B_i$ then $\forall i:\fun{f}{A}>\fun{f}{B_i}$.}
\lema{Acceptable programs are left terminating}{$P$ is acceptable if and only if $P$ is left terminating for all ground queries $\leftarrow A$.
\begin{proof}
\begin{description}
 \item [$\Rightarrow$] Assume $\leftarrow A$ is \inre{Left terminating}, let $\fun{f}{A}$ be the length of the longest derivation (under the computation rule) of $\leftarrow A$. Consider the ground clause $A\leftarrow B_1,\ldots,B_k$. If $B_1,\ldots,B_{i-1}$ succeeds, thus true in $M_P$, a model of $P$, then $\fun{f}{B_i}<\fun{f}{A}$. Thus $P$ is acceptable with respect to the \inre{Level mapping} $f$.
 \item [$\Leftarrow$] Assume an acceptable program $P$ and an infinite derivation $\leftarrow A,\ \leftarrow B_1,\ldots,B_k,\ ...$. By applying grounding on it there is a ground clause $A\leftarrow B_1\theta,\ldots,B_k\theta$. Since the derivation is infinite, there exists an $i$ such that $\leftarrow B_1\theta,\ldots,\leftarrow B_{i-1}\theta$ succeeds and $\leftarrow B_i\theta$ does not terminate. Thus $M_P\vDash B_1\theta,\ldots,B_{i-1}\theta$ and $M_P\subseteq I$ thus due to acceptability $\fun{f}{A}>\fun{f}{B_i\theta}$. By induction, one can construct an infinite decreasing sequence $\fun{f}{A}>\fun{f}{B_i}>\fun{f}{C_j\sigma}>\ldots$ of natural numbers, which is a contradiction.
\end{description}
\end{proof}}
\lema{Acceptable and left terminating for atomic queries}{If $P$ is \inre{Acceptable} with respect to \inre{Level mapping} $f$ and \inre{Model} $I$, then $P$ is \inre{Left terminating} for all atomic queries which are bounded with respect to $f$.
\begin{proof}
An infinite derivation for a bounded query can be instantiated into a infinite derivation for a ground query, which contradicts acceptability.
\end{proof}}
\lema{Atomic queries always terminate for acceptable programs}{If atomic queries bounded with respect to $f$ always \inre{Left terminate}, then $P$ is \inre{Acceptable} with respect to some \inre{Level mapping} $g$.
\begin{proof}
All ground queries are bounded with respect to $f$ hence \inre{Left terminate}. Thus $P$ is \inre{Acceptable} (according to the previous lemma).
\end{proof}}
\lema{Undecidability of termination}{No algorithm is possible for constructing the \inre{Level mapping}s.}
\subsection{Termination of Normal Programs}
\subsubsection{Strong termination}
\lema{Level mapping for normal programs}{Every level mapping for normal programs has the following property:
\begin{equation}
\fun{f}{\funm{not}{A}}=\fun{f}{A}
\end{equation}}
\defe{Acyclic program}{A program $P$ is acyclic if there exists a \inre{Level mapping} $f$ such that for every $A\leftarrow L_1,\ldots,L_n\in\funm{ground}{P}$, $\forall i:\fun{f}{A}>\fun{f}{L_i}$.}
\lema{Termination of ground queries}{SLDNF terminates on ground queries for acyclic programs.
\begin{proof}
Infinite derivation instantiates to an infinite ground derivation, this contradicts the acyclicity assumption.
\end{proof}}
\lema{Acyclicity of terminating programs}{SLDNF can terminate for a ground query while the program is not acyclic\footnote{Note that this is contrary to the SLD procedure for recurrency.}.
\begin{proof}
Assume the following program $P$: $\fun{p}{x}\leftarrow q,\funm{not}{\fun{p}{x}}$ which is not an acyclic program. When one introduces the query $\leftarrow\fun{p}{x}$, this fails finitely.
\end{proof}}
\lema{Termination of bounded queries}{If $P$ is acyclic with respect to \inre{Level mapping} $f$ and $L$ is bounded with respect to $f$, then SLDNF terminates for all queries $\leftarrow L$. Boundedness however does not imply success or failure.}
\lema{Properties of acyclic programs}{
\begin{enumerate}
 \item An \inre{Acyclic program} is an \inre{Inductive definition} and is \inre{Locally stratified}. It is the \inre{Level mapping} who divides $\funm{ground}{P}$ in \inre{Strata}.
 \item $T_P$ has a unique \inre{Fixpoint} $M_P$.
 \item $M_P$ is the perfect model of $T_P$.
 \item $M_P$ is the unique Herbrand model of $\funm{COMP}{P}$.
 \item $M_P$ is the unique model of $\funm{IND}{P}$.
 \item $\tupl{M_P,B_P\setminus M_P}$ is the unique \inre{Fixpoint} of $\Phi_P$.
 \item $M_P$ is the unique \inre{Well-founded model}.
 \item $M_P$ is the unique \inre{Stable model}.
\end{enumerate}}
\subsubsection{LDNF termination}
\lema{Acceptibility of normal programs}{Consider the rule $A\leftarrow B_1,\ldots,B_{j-1},\funm{not}{B_j},B_{j+1},\ldots,B_n\in\funm{ground}{P}$. If $I\vDash B_1,\ldots,B_{j-1},\funm{not}{B_j}$, is $\fun{f}{A}>\fun{f}{B_{j+1}}$? The \inre{Herbrand base} $B_P$ is a model. This gives: if $\mbox{true}\wedge\ldots\wedge\mbox{true}\wedge\mbox{false}$ then $\fun{A}{B_{j+1}}$ or if \mbox{false}, then $\fun{f}{A}>\fun{f}{B_{j+1}}$. Hence $B_{j+1}$ can be unbounded and non terminating. A solution can be to force $\fun{f}{A}>\fun{f}{B_j}\wedge\fun{f}{A}>\fun{f}{B_{j+1}}$.}
\lema{Requirements for the model to be used by acceptability}{
If $\funm{not}{B_k}$ succeeds, then $B_k$ is in the \inre{Finite failure set}. Thus $B_k$ does not belong to the \inre{Greatest fixpoint} of $T_P$ and $\funm{not}{B_k}$ is true in \inre{Model}s smaller or equal to the greatest \inre{Fixpoint} $I_{\vline B_k}\subseteq\funm{gfp}{T_P}_{\vline B_k}$. Positive attoms that succeed are part of $\funm{lfp}{T_P}$, they are true in any model larger or equal to $\funm{lfp}{T_P}$. Thus $\forall i:1\leq i\leq j:I_{\vline B_i}\supseteq\funm{lfp}{T_P}{|\vline B_i}$. If an atom has both positive and negative occurrences, one could use one model for the positive and one model for the negative occurrences.

}
\subsection{Framework for Automated Termination Analysis}
\subsubsection{Introduction}
\lema{Limitations of automation}{
There is no decision procedure. Hence the best is to try to find sufficient conditions for termination or non-termination. One cannot prove termination for all programs. Therefore searching for a level mapping allows one to prove termination. This however comes down to use the empty interpretation for the model of $B_k$ where negative literals are always true. It is fine however when \inre{Level mapping}s are based on the size of the data structures.}
\lema{Level mappings for automatic termination}{\inre{Level mapping}s are unnatural because they assign a \inre{Level mapping} to app predicates. Therefore \inre{Level mapping}s achieve two unrelated things: If $\fun{p}{t}$ recursively calls $\fun{p}{t'}$ then the SLD-tree for $\fun{p}{t'}$ is strictly smaller than the SLD-tree for $\fun{p}{t}$. If $\fun{p}{t}$ calls $\fun{q}{t'}$, then the SLD-tree for $\fun{q}{t'}$ is bounded.}
\note{Structure of a general framework}{A general framework should evidently focus on the loops of a problem: each time control reaches the same program point, there should be a decrease in the measure (element of a \inre{Well-founded set}) that captures the size of the program state. In logic program this means program points are calls to predicates, loops are calls to the same predicate. The framework should compute the calls of all atoms (or literals) which are selected for execution. By focusing on the recursion of the program: calls where the predicate already appears in the call-set. Then the framework should use a \inre{Level mapping} measuring the size of the calls. Note however that in this form of \inre{Level mapping}s, there is no interaction between \inre{Level mapping}s of different predicates.}
\subsubsection{LD-termination}
\defe{Collecting calls operator}{The \sub{} $T_P^C$ is a mapping of sets of (non-grounded) atoms to sets of atoms where we define $\fun{T_P^C}{S}$ as follows:
\begin{equation}
\fun{T_P^C}{S}=S\cup\accl{
\begin{array}{l|l}
&h\leftarrow b_1,\ldots,b_{i-1},b_i,\ldots\in P\wedge\\b_i\theta\sigma_1\ldots\sigma_{i-1}& a\in S\wedge\theta=\funm{mgu}{a,h}\wedge\\&\forall j:1\leq j<i:\sigma_j\mbox{ is a computed answer set of }b_j\theta\sigma_1\ldots\sigma_{j-1}
\end{array}}
\end{equation}}
\lema{Properties of the collecting calls operator}{\begin{enumerate}
 \item $T_P^C$ is \inre{Monotone}.
 \item $T_P^C$ is \inre{Continuous}.
 \item $T_P^C$ has a \inre{Least fixpoint}.
\end{enumerate}}
\defe{Calls operator}{We define the calls operator as:
\begin{equation}
\funm{Calls}{C,P}=\funm{lfp}{\fun{T_P^C}{C}}
\end{equation}
It is the set of all atoms that can be selected in an LD-derivation of a query $\leftarrow A$ for $A\in C$}
\lema{Left termination of the calls operator}{
A program $P$ is \inre{Left terminating} with respect to a set of \inre{Atomic class} $C$ if and only if it is \inre{Left terminating} for \funm{Calls}{C,P}.
}
\defe{Left recursive resultant}{By building a \inre{Partial LD-tree} for $\fun{p}{t}$ and terminate a branch when the selected atom is $\fun{p}{s}$. For each branch that represents a partial derivation extract the resultant $\fun{p}{t}\theta\leftarrow\fun{p}{s}\ldots$. $\theta$ is the substitution accumulated along the branch (and one can drop the $\ldots$ part).}
\defe{Level mapping for automated termination}{A \sub{} is a mapping $f$ from atoms to natural numbers where
\begin{equation}
\fun{f}{A}=\fun{\min}{\accl{\fun{f}{A\theta}|\theta\mbox{ is a grounding substitution}}}
\end{equation}. Therefore one can see a \sub{} as a \inre{Terminating condition} for a left recursive resultant $\fun{p}{t}\theta\leftarrow\fun{p}{s}$ if and only if $\fun{f}{\fun{p}{t}\theta}>\fun{f}{\fun{p}{s}}$.}
\defe{Rigid level mapping}{A \inre{Level mapping} is \inre{Rigid} if $\fun{f}{\fun{p}{t}\theta}=\fun{f}{\fun{p}{t}\theta}$ for all $\theta$.}
\defe{Acceptable program with respect to a set of calls $C$}{A program is \inre{Acceptable} with respect to a set of calls $C$ if there is a \inre{Level mapping} for $f$ for each recursive predicate such that: for each call $\fun{p}{t}\in C$ and for each \inre{Left recursive resultant} $\fun{p}{t}\theta\leftarrow\fun{p}{s}$, $\fun{f}{\fun{p}{t}}>\fun{f}{\fun{p}{s}}$. If $f$ is rigid for each call in $C$ than $f$ is a terminating condition for each left recursive resultant.}
\subsubsection{Binary unfolding}
\defe{Binary unfolding operator}{A \sub{} $T_P^b$ maps binary clauses to binary clauses such that:
\begin{equation}
\fun{T_P^b}{B}=\accl{
\begin{array}{l|l}
\brak{h\leftarrow b}\theta&\brak{h\leftarrow b_1,\ldots,b_m\in P}\wedge\funm{condition}{B,P,h,b_1,\ldots,b_m,\theta}\\
\end{array}}
\end{equation}
There are three ways to satisfy this condition:
\begin{eqnarray}
\exists i:1\leq i\leq m:\accl{h_1\leftarrow\mbox{true},\ldots,h_{i-1}\leftarrow\mbox{true}}\subseteq B\wedge\theta=\funm{mgu}{\tupl{h_1,\ldots,h_{i-1}},\tupl{b_1,\ldots,b_{i-1}}}\wedge b=b_i\\
\exists i:1\leq i\leq m:\accl{h_1\leftarrow\mbox{true},\ldots,h_{i-1}\leftarrow\mbox{true},h_i\leftarrow c}\subseteq B\wedge c\neq\mbox{true}\wedge\theta=\funm{mgu}{\tupl{h_1,\ldots,h_i},\tupl{b_1,\ldots,b_i}}\wedge b=c\\
\exists i:1\leq i\leq m:\accl{h_1\leftarrow\mbox{true},\ldots,h_m\leftarrow\mbox{true}}\subseteq B\wedge\theta=\funm{mgu}{\tupl{h_1,\ldots,h_m},\tupl{b_1,\ldots,b_m}}\wedge b=\mbox{true}
\end{eqnarray}}
\defe{Binary unfolding of a program}{The \sub{} $P$ is the \inre{Least fixpoint} of the \inre{Binary unfolding operator} $T_P^b$, thus:
\begin{equation}
\fun{bin\_unf}{P}=\fun{lfp}{T_P^b}
\end{equation}}
\lema{Properties of the binary unfolding of a program}{The \inre{Binary unfolding} is goal dependent: it depends on a particular query, it is \inre{Closed under unfolding}: if $\fun{p}{x}\leftarrow\fun{p}{y}\in\funm{bin\_unf}{P}$ and $\fun{p}{y'}\leftarrow\fun{p}{z}\in\funm{bin\_unf}{P}$ and $\theta=\funm{mgu}{\fun{p}{y},\fun{p}{y'}}$ then $\brak{\fun{p}{x}\leftarrow\fun{p}{z}}\theta\in\fun{bin\_unf}{P}$. \inre{Binary unfolding} also allows to computed all calls and answers selected during an \inre{LD-derivation} of an atomic call $\leftarrow c$ or a set of atomic class $C$.}
\subsubsection{Rounding up}
\subsection{Summary}